HSC Synthetic Pipeline I: Basic Design and Characterization of HSC Photometric Performance

### 2017-04-20 draft/ms_v3.tex
Quickly go through Susan's edits

    \song{Song Huang's edit}
    For star selection, we use the sample from \citet{Leauthaud2007}.
    We use not only the \texttt{SExtractor} stellar classification index
    (\texttt{CLASS\_STAR}) but also the peak surface brightness above background
    (\texttt{MU\_MAX}) and half-light radius (\texttt{RHL}) to select stars with clean
    photometry.
    The classification is reliable down to $I_{\mathrm{F814W}}{\sim}25.2$ mag.
    We match this sample of stars with the HSC UltraDeep--COSMOS data that currently
    reach to about ${\sim}27.2$ mag in $i$-band.
    After filtering the matched sample through basic quality cuts (see \ref{app:qc}),
    we gather five-band HSC photometry for 14,472 stars.
    To get a larger sample for the test, we resample the five-band magnitude
    distributions using \texttt{astroML} (\citealt{astroml}) alternative
    implementation of the \texttt{extreme-deconvolution} algorithm developed by
    \citet{Bovy2011}
    \footnote{\texttt{https://github.com/jobovy/extreme-deconvolution/}} to
    generate 100,000 synthetic stars.
    In Fig \ref{fig:star_sample}, we show that the magnitude and color distributions of
    the real COSMOS stars are nicely recovered by the resampled stars in all
    five-band magnitude and two color-color space distributions.

@@

    We match the results with the input catalogs using a 2--pixel matching radius
    to generate output catalogs for \forced{} and \unforced{} photometry
    in each band.
    We also match the input catalogs with the original HSC catalogs to identify the
    ``ambiguously blended'' objects that are right on top of the flux peak of a real
    object. \susan{Is ambiguously blended well-understood by your reader? Or is the definition that they are objects that are right on top of the flux peak of a real object? If so, change sentence to the following: We also match the input catalogs with the original HSC catalogs to identify the  ``ambiguously blended'' objects, which are defined as those objects that are right on top of the flux peak of a real
    object.}
    The ambiguously blended objects are useful in the WL-related statistical tests (see Murata et al.
    in prep. for more details); but since they will not help test photometry, they will be excluded later.

@@

    Here we discuss \hscpipe{}'s general performance, assessed mainly through  \forced{} PSF photometry (for stars) and (\cmodel{}) photometry (for galaxies).
 Although  \hscpipe{} does provide other options (e.g., aperture and Kron photometry),
these two are the only options that consider the effects of PSF in different bands and are consistent across all bands in terms of position and shape. Therefore, they are the most appropriate and relevant to star and galaxy magnitudes and colors.

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic stars and the $log(S/N)$ measured by \hscpipe{} PSF
        photometry.
        Hexagonal binned density plots and contours show the distributions for
        stars in both the good- and bad-seeing \tracts{}.
        The highly blended stars are highlighted using scatter plots.
        The gray dashed line marks $S/N = 5$. \susan{"grey" is an acceptable spelling; however, gray is more commonly used in the United States.}
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    In \hscpipe{}, the PSF magnitude is derived using a matched-filter method that
    depends on the best-fit PSF model and uses the Lanczos interpolation to shift
    the PSF model.
    The error of the PSF magnitude from \hscpipe{} only considers the per-pixel noise,
    and does not include centroid uncertainty.
    \hscpipe{} also estimates aperture correction for PSF photometry.
    For more details please see Bosch\etal (in prep.). \susan{Suggest you not include a section number for a document in preparation; section numbers can change before a text is published. }

    In our tests, we randomly inject ${\sim}100,000$ stars into each \tract{} in all
    five bands. Approximately \susan{Spell out approximately and delete approximate sign before 3-4.}
 ${\sim}3$--$4$\% of the stars are located within 2 pixels of the centroids
    of real objects; we remove those stars from the sample to avoid confusion in the comparisons.
    After this step, ${\sim}1$--$3$\% of the remaining synthetic stars still have no match in the \synpipe{} outputs.
    They could either be below the detection threshold or ambiguously blended with real objects.
    The input $i$-band magnitudes of these unmatched stars are on average ${\sim}0.9$
    mag fainter than the matched ones.
    The fraction of unmatched stars is slightly higher in the bad seeing \tract{}; this result is consistent with our expectation that magnitude depends on seeing conditions,
    although the difference is not very significant.
    For the matched stars, we select the primary detections
    (\texttt{detect.is-primary=True}) that have good photometric quality
    (see Appendix \ref{app:qc} for details).
    This gives us ${\sim}83,000$ stars in each \tract{}.

@@

    In the following comparisons, we show the results from the
    \textbf{good-seeing} and \textbf{bad-seeing} \tracts{} separately because it
    is important to understand how the seeing condition affects photometric accuracy as well as to test whether \hscpipe{} can deliver unbiased photometry under different  seeing conditions. \susan{Later you use "slightly worse" and "better" seeing; if you want to use those terms, perhaps define here where you make the distinction? For example, can you quantify slightly worse and better?}

    Besides seeing, ``blendedness'' ($b$; parameter that describes on what level
    the flux distribution of one object is affected by others; please see
    Appendix \ref{app:defineb} for precise definition, and see Murata\etal in prep.
    for details) is another factor that influences the photometric accuracy.
    Therefore, we separate the samples into relatively isolated and highly blended
    groups based on the $b$ values.
    Assuming zero pixel noise, $b$ of object $A$ reflects the fraction of flux
    within the  $A$ \texttt{Footprint} that comes from other objects.
    Ideally, $b=0$, which means perfect isolation; $b{\sim}1$ means completely
    confused with another object.
    Here we use $b>0.05$ to define highly \susan{Later you say that $b>0.05$ is "highly blended"; it needs to be consistent. If inserting "highly" is incorrect here, please define b for highly blended stars since you use highly blended throughout the text.} blended stars (${\sim}4$--$9$\% are blended).
    The fraction of highly blended stars slightly increases in the bad seeing \tract{}.
    We highlight these stars separately as well, and will discuss more about the
    impact of blendedness in Section \ref{ssec:blendedness}.

@@

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$- and
    $i$-bands at average seeing conditions, which is consistent with the values found
    in \citet{HSCDR1}. \susan{Is there a value for "average seeing conditions"?}
    At the same time, it is worth reminding HSC data users that the detection
    limit will show spatial variations due to the seeing conditions.

@@

    For relatively isolated stars, \hscpipe{} provides unbiased PSF photometry down to a
    very faint magnitude in $i$- and $z$-bands.
    For $g$-, $r$-, and $y$-bands, we find small offsets in the distributions of magnitude
    differences, which indicates \hscpipe{} tends to underestimate the fluxes of stars
    in these bands by ${\sim}0.01-0.02$ mag.
    Although the exact cause of this offset is unclear (it could be related to
    differences in seeing), it will not affect the accuracies of magnitudes of \susan{should "of" be "or"?} colors.

@@

    We further illustrate the accuracies of PSF colors via comparing the distributions
    of colors for the input synthetic stars and the stars recovered by \hscpipe{}
    (see Fig \ref{fig:psf_cdist}).
    Using four different colors and two \susan{2-D?} color--color planes, we show that the \forced{}
    PSF colors can accurately reproduce the distributions of all four colors and the
    2-D \susan{or two?} color-color distributions.

@@

  For galaxies, \hscpipe{} uses a \cmodel{} photometry algorithm
     that is an improved version of the SDSS \cmodel{}
    photometry.
    For more details, please see Bosch\etal (in prep.).\susan{Suggest you not include a section number for a document in preparation; section numbers can change before a text is published. }
    In short, \cmodel{} tries to fit each object using a combination of de Vaucouleurs
    and exponential profiles that are approximated by multiple Gaussian profiles.
    Despite the limitation of the \cmodel{} method (e.g., sensitive to background
    subtraction and deblender failure), it can deliver robust PSF-corrected fluxes and
    colors for galaxies.

@@

    We also use $b>0.1$ to define samples of highly blended galaxies.
    Due to the extended nature of the galaxies, the $b$ distributions for synthetic galaxies are
    very different from the  $b$ distributions for synthetic stars, and they also skew toward higher values (especially for the
    bad-seeing \tract{}).
    About ${\sim}7$--$9$\% \susan{Delete the approximately symbol when you use "About"; it's redundant.} of synthetic galaxies are highly blended by this standard.
    The fraction is only slightly higher in the bad-seeing \tract{}, and the
    $g$-band has the highest fraction of highly blended galaxies.
    We highlight the highly blended ones in comparisons, and there are other relevant
    discussions in Section \ref{ssec:blendedness}.

@@

    Similar to the synthetic stars, the worse the seeing condition, the more
    we noticed a decrease in \s2n{} at the same magnitude (see the figure \susan{cite actual figure number.} for
    $i$-band).
    Meanwhile, the highly blended synthetic galaxies do not show significantly
    different \s2n{} distributions here.

@@

    Star--galaxy separation is an important aspect of a cosmology survey,
    and it becomes increasingly difficult at very faint magnitude ($i<24.0$ mag).
    In \hscpipe{}, the star--galaxy classification primarily depends on the extendedness
    value, which is measured by the magnitude difference between PSF and \cmodel{}
    photometry in $i$-band.
    For more details about this algorithm, please see Bosch\etal
    (in prep.). \susan{Suggest you not include a section number for a document in preparation; section numbers can change before a text is published. }

@@

        \item High blendedness ($b>0.1$) \susan{This is the value you've given for highly blended galaxies; should the text in the parentheses read ($b>0.1$ for galaxies and $b>0.05$ for stars)}impacts the accuracy of both PSF and \cmodel{}
            photometry, especially for stars.
            For highly blended stars, the \forced{} PSF photometry on average
            overestimates the magnitudes of stars by 0.1–-0.2 mag.
            For galaxies, high-blendedness typically adds an additional 0.05 mag
            uncertainty in both magnitude and color estimates.

@@

%% ------------------------------------------------------------------------------------ %%
    %% Other applications
    Although we mainly focus on the general behaviors of photometric measurements in this study,
    \synpipe{} is quite flexible and has many other applications for any HSC
    dataset.
    For instance,  Murata\etal (in prep.) uses \synpipe{} to test
    the impacts of extremely blended \susan{Quantify? Is "extremely blended" more blended than "highly blended"?} objects on the WL measurements.
    Also, \citep{Niikura2017} applies \synpipe{} to the high-cadence M31 HSC observations
    to help users search for interesting microlensing events.
  \synpipe{} is also currently being used to estimate the completeness of the selections
    for high-$z$ LBG and Lyman- $\alpha$ emitters
    (Ono\etal in prep.; Konno\etal in prep.), check the dependence of the detection and
    completeness of high-$z$ LBG pairs on the pair separation (Harikane\etal in prep.),
    identify low surface--brightness dwarf galaxies (Greco\etal in prep.),
    and test photometric and WL measurements around nearby clusters (Chiu\etal in prep.).

@@

        \item  \synpipe{} simply adopts the PSF measured by \hscpipe{} and
            uses it as a model of point source and PSF convolution for a galaxy.
            Hence, it cannot be used to test how the uncertainty of the PSF modeling
            affects the photometry and shape measurements.
            This could be important for regions with exquisite seeing \susan{Quantify "exquisite"? Is it better than "good" or "better" seeing? Suggest using consistent adjectives.}  that makes the
            PSF modeling difficult, and for accurate WL measurements.

@@

    % Software part
    This research made use of
    {\texttt{Astropy}},
        a community-developed core Python package for Astronomy (\citealt{Astropy};
        \url{http://www.astropy.org/});
    {\texttt{astroML}},
        a machine learning library for astrophysics (\citealt{astroml};
        \url{http://www.astroml.org/});
    {\texttt{SciPy}},
        an open source scientific tool for Python (\citealt{SciPy};
        \url{http://www.scipy.org/});
    {\texttt{NumPy}},
        a fundamental package for scientific computing with Python (\citealt{NumPy};
        \url{http://www.numpy.org/});
    {\texttt{Matplotlib}},
        a 2-D plotting library for Python (\citealt{Matplotlib};
        \url{http://matplotlib.org/}); and
    {\texttt{scikit-learn}},
        a machine learning library in Python (\citealt{scikit-learn};
        \url{http://scikit_learn.org/} \susan{I think correct URL is http://scikit-learn.org/ (hyphen, not underscore.}).

@@

%% ------------------------------------------------------------------------------------ %%
\susan{Susan did not review bibliographic citations.}
\begin{thebibliography}{}
    \label{sec:ref}
    \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

>>>>

    For star selection, we use the sample from \citet{Leauthaud2007}.
    We use not only the \texttt{SExtractor} stellar classification index
    (\texttt{CLASS\_STAR}) but also the peak surface brightness above background
    (\texttt{MU\_MAX}) and half-light radius (\texttt{RHL}) to select stars with clean
    photometry.
    The classification is reliable down to $I_{\mathrm{F814W}}{\sim}25.2$ mag.
    We match this sample of stars with the HSC UltraDeep--COSMOS data that currently
    reach to about ${\sim}27.2$ mag in $i$-band.
    After filtering the matched sample through basic quality cuts (see \ref{app:qc}),
    we gather five-band HSC photometry for 14,472 stars.
    To get a larger sample for the test, we resample the five-band magnitude
    distributions using \texttt{astroML} (\citealt{astroml}) alternative
    implementation of the \texttt{extreme-deconvolution} algorithm developed by
    \citet{Bovy2011}
    \footnote{\texttt{https://github.com/jobovy/extreme-deconvolution/}} to
    generate 100,000 synthetic stars.
    In Fig \ref{fig:star_sample}, we show that the magnitude and color distributions of
    the real COSMOS stars are nicely recovered by the resampled stars in all
    five-band magnitude and two color-color space distributions.

@@

    We match the results with the input catalogs using a 2--pixel matching radius
    to generate output catalogs for \forced{} and \unforced{} photometry
    in each band.
    We also match the input catalogs with the original HSC catalogs to identify the
    ``ambiguously blended'' objects, which are defined as so extreme blends that 
    some objects are detected as a single object.
    The ambiguously blended objects are useful in the WL-related statistical tests 
    (see Murata et al. in prep. for more details); but since they will not help test 
    photometry, they will be excluded later.

@@

    Here we discuss \hscpipe{}'s general performance, assessed mainly through \forced{} 
    PSF photometry (for stars) and (\cmodel{}) photometry (for galaxies).
    Although \hscpipe{} does provide other options (e.g., aperture and Kron photometry),
    these two are the only options that consider the effects of PSF in different bands 
    and are consistent across all bands in terms of position and shape. T
    Therefore, they are the most appropriate and relevant methods for star and galaxy.
    \song{Edit stops here}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic stars and the $log(S/N)$ measured by \hscpipe{} PSF
        photometry.
        Hexagonal binned density plots and contours show the distributions for
        stars in both the good- and bad-seeing \tracts{}.
        The highly blended stars are highlighted using scatter plots.
        The gray dashed line marks $S/N = 5$.
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    In \hscpipe{}, the PSF magnitude is derived using a matched-filter method that
    depends on the best-fit PSF model and uses the Lanczos interpolation to shift
    the PSF model.
    The error of the PSF magnitude from \hscpipe{} only considers the per-pixel noise,
    and does not include centroid uncertainty.
    \hscpipe{} also estimates aperture correction for PSF photometry.
    For more details please see Bosch\etal (in prep.).

    In our tests, we randomly inject ${\sim}100,000$ stars into each \tract{} in all
    five bands. Approximately ${\sim}3$--$4$\% of the stars are located within 2 pixels of the centroids
    of real objects; we remove those stars from the sample to avoid confusion in the comparisons.
    After this step, ${\sim}1$--$3$\% of the remaining synthetic stars still have no match in the \synpipe{} outputs.
    They could either be below the detection threshold or ambiguously blended with real objects.
    The input $i$-band magnitudes of these unmatched stars are on average ${\sim}0.9$
    mag fainter than the matched ones.
    The fraction of unmatched stars is slightly higher in the bad seeing \tract{}; this result is consistent with our expectation that magnitude depends on seeing conditions,
    although the difference is not very significant.
    For the matched stars, we select the primary detections
    (\texttt{detect.is-primary=True}) that have good photometric quality
    (see Appendix \ref{app:qc} for details).
    This gives us ${\sim}83,000$ stars in each \tract{}.

@@

    In the following comparisons, we show the results from the
    \textbf{good-seeing} and \textbf{bad-seeing} \tracts{} separately because it
    is important to understand how the seeing condition affects photometric accuracy as well as to test whether \hscpipe{} can deliver unbiased photometry under different  seeing conditions.

    Besides seeing, ``blendedness'' ($b$; parameter that describes on what level
    the flux distribution of one object is affected by others; please see
    Appendix \ref{app:defineb} for precise definition, and see Murata\etal in prep.
    for details) is another factor that influences the photometric accuracy.
    Therefore, we separate the samples into relatively isolated and highly blended
    groups based on the $b$ values.
    Assuming zero pixel noise, $b$ of object $A$ reflects the fraction of flux
    within the  $A$ \texttt{Footprint} that comes from other objects.
    Ideally, $b=0$, which means perfect isolation; $b{\sim}1$ means completely
    confused with another object.
    Here we use $b>0.05$ to define highly blended stars (${\sim}4$--$9$\% are blended).
    The fraction of highly blended stars slightly increases in the bad seeing \tract{}.
    We highlight these stars separately as well, and will discuss more about the
    impact of blendedness in Section \ref{ssec:blendedness}.

@@

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$- and
    $i$-bands, which is consistent with the values found
    in \citet{HSCDR1} for average seeing conditions.
    At the same time, it is worth reminding HSC data users that the detection
    limit will show spatial variations due to the seeing conditions.

@@

    For relatively isolated stars, \hscpipe{} provides unbiased PSF photometry down to a
    very faint magnitude in $i$- and $z$-bands.
    For $g$-, $r$-, and $y$-bands, we find small offsets in the distributions of magnitude
    differences, which indicates \hscpipe{} tends to underestimate the fluxes of stars
    in these bands by ${\sim}0.01-0.02$ mag.
    Although the exact cause of this offset is unclear (it could be related to
    differences in seeing), it will not affect the accuracies of magnitudes or colors.

@@

    We further illustrate the accuracies of PSF colors via comparing the distributions
    of colors for the input synthetic stars and the stars recovered by \hscpipe{}
    (see Fig \ref{fig:psf_cdist}).
    Using four different colors and two 2--D color--color planes, we show that the \forced{}
    PSF colors can accurately reproduce the distributions of all four colors and the
    2-D color-color distributions.

@@

  For galaxies, \hscpipe{} uses a \cmodel{} photometry algorithm
     that is an improved version of the SDSS \cmodel{}
    photometry.
    For more details, please see Bosch\etal (in prep.).
    In short, \cmodel{} tries to fit each object using a combination of de Vaucouleurs
    and exponential profiles that are approximated by multiple Gaussian profiles.
    Despite the limitation of the \cmodel{} method (e.g., sensitive to background
    subtraction and deblender failure), it can deliver robust PSF-corrected fluxes and
    colors for galaxies.

@@

    We also use $b>0.1$ to define samples of highly blended galaxies.
    Due to the extended nature of the galaxies, the $b$ distributions for synthetic galaxies are
    very different from the  $b$ distributions for synthetic stars, and they also skew toward higher values (especially for the
    bad-seeing \tract{}).
    About $7$--$9$\% of synthetic galaxies are highly blended by this standard.
    The fraction is only slightly higher in the bad-seeing \tract{}, and the
    $g$-band has the highest fraction of highly blended galaxies.
    We highlight the highly blended ones in comparisons, and there are other relevant
    discussions in Section \ref{ssec:blendedness}.

@@

    Similar to the synthetic stars, the worse the seeing condition, the more
    we noticed a decrease in \s2n{} at the same magnitude (see Fig \ref{fig:psf_s2n}) 
    for $i$-band).
    Meanwhile, the highly blended synthetic galaxies do not show significantly
    different \s2n{} distributions here.

@@

    Star--galaxy separation is an important aspect of a cosmology survey,
    and it becomes increasingly difficult at very faint magnitude ($i<24.0$ mag).
    In \hscpipe{}, the star--galaxy classification primarily depends on the extendedness
    value, which is measured by the magnitude difference between PSF and \cmodel{}
    photometry in $i$-band.
    For more details about this algorithm, please see Bosch\etal
    (in prep.).

@@

        \item High blendedness ($b>0.1$) impacts the accuracy of both PSF and \cmodel{}
            photometry, especially for stars.
            For highly blended stars, the \forced{} PSF photometry on average
            overestimates the magnitudes of stars by 0.1–-0.2 mag.
            For galaxies, high-blendedness typically adds an additional 0.05 mag
            uncertainty in both magnitude and color estimates.

@@

%% ------------------------------------------------------------------------------------ %%
    %% Other applications
    Although we mainly focus on the general behaviors of photometric measurements in this study,
    \synpipe{} is quite flexible and has many other applications for any HSC
    dataset.
    For instance,  Murata\etal (in prep.) uses \synpipe{} to test
    the impacts of the ambiguously blended objects on the WL measurements.
    Also, \citep{Niikura2017} applies \synpipe{} to the high-cadence M31 HSC observations
    to help users search for interesting microlensing events.
  \synpipe{} is also currently being used to estimate the completeness of the selections
    for high-$z$ LBG and Lyman- $\alpha$ emitters
    (Ono\etal in prep.; Konno\etal in prep.), check the dependence of the detection and
    completeness of high-$z$ LBG pairs on the pair separation (Harikane\etal in prep.),
    identify low surface--brightness dwarf galaxies (Greco\etal in prep.),
    and test photometric and WL measurements around nearby clusters (Chiu\etal in prep.).

@@

        \item  \synpipe{} simply adopts the PSF measured by \hscpipe{} and
            uses it as a model of point source and PSF convolution for a galaxy.
            Hence, it cannot be used to test how the uncertainty of the PSF modeling
            affects the photometry and shape measurements.
            This could be important for regions with exquisite seeing 
            (e.g. FWHM${\sim}0.4$\asec{}) that makes the
            PSF modeling difficult, and for accurate WL measurements.

@@

    % Software part
    This research made use of
    {\texttt{Astropy}},
        a community-developed core Python package for Astronomy (\citealt{Astropy};
        \url{http://www.astropy.org/});
    {\texttt{astroML}},
        a machine learning library for astrophysics (\citealt{astroml};
        \url{http://www.astroml.org/});
    {\texttt{SciPy}},
        an open source scientific tool for Python (\citealt{SciPy};
        \url{http://www.scipy.org/});
    {\texttt{NumPy}},
        a fundamental package for scientific computing with Python (\citealt{NumPy};
        \url{http://www.numpy.org/});
    {\texttt{Matplotlib}},
        a 2-D plotting library for Python (\citealt{Matplotlib};
        \url{http://matplotlib.org/}); and
    {\texttt{scikit-learn}},
        a machine learning library in Python (\citealt{scikit-learn};
        \texttt{http://scikit-learn.org/}.

@@

%% ------------------------------------------------------------------------------------ %%
\begin{thebibliography}{}
    \label{sec:ref}
    \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
